---
output: github_document
---

<!-- 
README.md is generated from README.Rmd. 
Please edit that file 
-->

```{r include=FALSE}
# Create .nojekyll file in the /docs folder
file.create('docs/.nojekyll')
# Avoid problems with "Onedrive - Universidade da Coruña" 
# devtools::check(getwd()) 
# pkgdown::build_site()
```

# npfda: Nonparametric functional data analysis

### `r paste("Version", packageVersion("npfda"))`

This package implements nonparametric methods for inference on (multidimensional) functional data using the tools available in package [`npsp`](https://rubenfcasal.github.io/npsp).


```{r echo=FALSE, results='asis'}
if ('github_document' == rmarkdown::all_output_formats(knitr::current_input()) && !knitr:::is_html_output()) 
   cat('For more information visit the [web page](https://rubenfcasal.github.io/npfda) of the package.')
```


## Main functions

*   `npf.data()` defines a (multidimensional) functional data set.

Nonparametric methods for inference on the functional trend, the functional variance 
and the variogram:

*  `npf.fit()` (NOT YET IMPLEMENTED) (automatically) fits a nonparametric functional model 
    by estimating the trend, the conditional variance and the variogram.

*  `locpol()`, `np.var()` and `np.svar()` methods use local polynomial kernel 
   smoothing to compute nonparametric estimates of the functional trend, the functional 
   variance and the (intra-curve) variogram (or their first derivatives), respectively. 


See the [Reference](https://rubenfcasal.github.io/npfda/reference/index.html) for the complete list of functions.


## Installation

`npfda` is not available from CRAN, but you can install the development version from github with:

```{r install, eval=FALSE}
# install.packages("remotes")
remotes::install_github("rubenfcasal/npfda")
```

<!-- 
Note also that, as this package requires compilation, Windows users need to have previously installed the appropriate version of [Rtools](https://cran.r-project.org/bin/windows/Rtools/), and OS X users need to have installed [Xcode](https://apps.apple.com/us/app/xcode/id497799835).

Alternatively, Windows users may install the corresponding *npfda_X.Y.Z.zip* file in the [releases section](https://github.com/rubenfcasal/npfda/releases/latest) of the github repository.

For R versions 4.2.x under Windows: 
``` r
install.packages('https://github.com/rubenfcasal/npfda/releases/download/v0.1-1/npfda_0.1-1.zip',
                 repos = NULL)
``` 
-->

## Authors

* [Rubén Fernández-Casal](https://rubenfcasal.github.io) (Dep. Mathematics, University of A Coruña, Spain).

* Sergio Castillo-Páez (Universidad de las Fuerzas Armadas ESPE, Ecuador).

* Miguel Flores (Faculty of Administrative Sciences, Escuela Politécnica Nacional, Ecuador).

Please send comments, error reports or suggestions to [rubenfcasal@gmail.com](mailto:rubenfcasal@gmail.com).


## Acknowledgments

This research has been supported by MICINN (Grant PID2020-113578RB-I00). 
The research of Rubén Fernández-Casal has been supported by the Xunta de Galicia (Grupos de Referencia Competitiva ED431C-2020-14 and Centro de Investigación del Sistema Universitario de Galicia ED431G 2019/01). All these grants were co-financed by the ERDF. 
The research of Sergio Castillo Páez has been supported by the Universidad de las Fuerzas Armadas ESPE, from Ecuador.


## References

* Fernández-Casal R. (2023) npsp: Nonparametric spatial (geo)statistics. R package version 0.7-11, *https://rubenfcasal.github.io/npsp*.

* Fernández-Casal R., Castillo-Páez S. and García-Soidán P. (2017), Nonparametric estimation of the small-scale variability of heteroscedastic spatial processes, *Spa. Sta.*, **22**, 358-370, [DOI](https://doi.org/10.1016/j.spasta.2017.04.001).

* Shapiro A. and Botha J.D. (1991) Variogram fitting with a general class of conditionally non-negative definite functions. *Computational Statistics and Data Analysis*, **11**, 87-96.

